Enhancing Forecast Reconciliation: A Study of Alternative Covariance Estimators

### 2025-09-23 literature/paper_draft.qmd
revise tourism result discussion

---
title: "Enhancing Forecast Reconciliation: A Study of Alternative Covariance Estimators"
author: 
  - Vincent Su
  - Shanika Wickramasuriya (supv.)
  - George Athanasopoulos (supv.)
format: 
  pdf:
    include-in-header:
      text: |
        <!-- \usepackage{amsmath} -->
        <!-- \usepackage{mathspec} -->
        \usepackage{algorithm}
        \usepackage{algpseudocode}
        \usepackage[colorlinks=true, linkcolor=blue, citecolor=blue, urlcolor=blue]{hyperref}
        \usepackage{array}
    mathspec: true
    number-sections: true
# crossref:
#   custom:
#     - kind: float
#       key: alg
#       reference-prefix: Algorithm
#       caption-prefix: Algorithm
#       # for PDF output only, tell LaTeX to use the `algorithm` env:
#       latex-env: algorithm
execute:
  cache: true
  echo: false
  warning: false
bibliography: D:/Github/Recon_Honours_Thesis/references.bib
# bibliographystyle: apa
csl: D:/Github/Recon_Honours_Thesis/apa.csl
editor_options: 
  chunk_output_type: console
---

@@

```{r library}
library(ggplot2)
library(patchwork)
library(dplyr)
library(tidyr)
library(fable)
library(fabletools)
library(feasts)
library(tsibble)
library(lubridate)
library(knitr)
library(kableExtra)
library(stringr)

@@

# Abstract {-}

@@

<!-- -   Talk about description of the hierarchical time series data set (e.g., economic, financial, or synthetic data). -->
<!-- -   Characteristics such as dimensionality, frequency, and hierarchical structure. -->

@@

<!-- -   Design different cases of simulation studies and real-data experiments. -->
<!-- -   Metrics: Forecast accuracy (e.g., RMSE, MAE), reconciliation error reduction, and matrix stability... -->
<!-- -   Visualisation... -->

@@

<!-- Rather than collapsing each group into a single aggregate as in previous practice, we impose a more intricate, multi‐level aggregation path to stress‐test reconciliation methods. We first form ten intermediate series by summing bottom‐level series in consecutive blocks of ten. These ten series are then assigned to three second‐level aggregates--groups of four, three, and four--and finally summed to produce a single top‐level series. This asymmetric hierarchy creates overlapping correlation patterns: some level‐2 series share bottom‐level groups, while others draw from both. -->
<!-- emulating practical scenarios such as regional sales aggregations that span multiple product categories or overlapping territories. -->

@@

@@

```{r}
library(stringr)
desired_order <- c(
  "Australia", "Australia_by_Purpose",
  "State",     "State_by_Purpose",
  "Zone",      "Zone_by_Purpose",
  "Region",    "Region_by_Purpose"
)
```

```{r fig tourism results}
#| label: fig-tourism-results
#| fig-cap: "Relative improvement of the MSE of reconciled forecasts over the base forecasts for the Australian domestic tourism flows"
#| fig.width: 8
#| fig.height: 10
#| out.width: "100%"

tourism_MSE <- readRDS("./literature/lib/tourism_MSE.rds")

# % improv plot by level
tourism_MSE |>
  group_by(series, h) |>
  mutate(
    base_MSE = MSE[.model == "base"]
  ) |>
  ungroup() |>
  group_by(.model, level, h) |>
  summarise(
    MSE = mean(MSE),
    base_MSE = mean(base_MSE),
    pct_change = (MSE - base_MSE) / base_MSE * 100
  ) |>
  filter(.model != "mint_sample") |>
  # Reorder the levels according to the desired order
  mutate(
    level = factor(
      level,
      levels = desired_order,
      labels = str_replace_all(desired_order, "_", " ")  # replace "_" with " "
    )
  ) |> 
  ggplot(aes(x = h, y = pct_change, color = .model)) +
    geom_line() +
    labs(x = "Horizon", y = "% improvements") +
    facet_wrap(~ level, ncol=2, scales = "free_y") +
    theme_bw() +
    scale_x_continuous(limits = c(1, 12), expand = c(0, 0)) +
    theme(legend.position = "bottom", legend.title = element_blank())
```

The data are monthly time series spanning from January 1998 to December 2019, resulting in 264 observations per series, producing a canonical "$n \ll p$" setting which is ideal for evaluating reconciliation approaches that rely on high-dimensional covariance estimation. The extreme dimensionality over sample size mirrors many contemporary business problems, for instance, Starbucks drink sales. Tourism demand is also economically vital yet highly volatile, with geographical and purpose‑specific patterns create a realistic stress‑test for reconciliation algorithms. Consequently, this panel offers both a rich policy case study and a stringent statistical laboratory for comparing reconciliation strategies that exploit cross-series information to improve forecasts when historical data are scarce.

@@

@Wickramasuriya2019-xq also argued that modelling spatial autocorrelations would be challenging as in the case of a large collection of time series. Reconciliation approaches have the advantage to implicitly model this spatial autocorreltion structure, especially when the MinT method is used.

To assess forecasting performance, we adopt a rolling-window cross-validation scheme. Beginning with the first 96 monthly observations (January 1998-December 2005) as the initial training set, we obtain the best-fitted ARIMA model for each of the 560 series via the automatic algorithm in the fabletools package, by minimising AICc [@O-Hara-Wild2024-we]. The 1- to 12-step-ahead base forecasts are then generated by these ARIMA models. These incoherent base forecasts are reconciled via OLS, MinT with shrinkage (mint_shr), and MinT with the NOVELIST estimator (mint_n). We then roll the training window forward by one month and refit all models, rebuild reconciliations, and produce another batch of 1- to 12-step-ahead forecasts, repeating until December 2018. In total, this results in 156 out-of-sample windows and an equal number of forecast sets for each series.

Note that the number of series is larger than the number of observations (560 compared to 96), hence the sample covariance matrix is not positive definite and will not be considered.

The results are presented in @fig-tourism-results, which shows the relative improvement of the MSE of reconciled forecasts over the base forecasts for the Australian domestic tourism flows. The results are grouped by levels of aggregation. MinT variants show improvement over the base ARIMA forecasts for middle to lower aggregation levels. We hardly differentiate the point accuracy performance between mint_shr and mint_n in these level, except at the most disaggregated, where mint_shr slightly edges out mint_n.

```{r fig STL tourism}
#| label: fig-STL-tourism
#| fig-cap: "Monthly domestic overnight tourism nights in Australia by state, with STL decomposition to extract trend component"
#| fig.width: 8
#| fig.height: 6
#| out.width: "100%"
#| fig.align: "center"

visnights_full <- readRDS("literature/lib/visnights_full.rds")

visnights_full |> 
  filter(
    is_aggregated(Purpose) & is_aggregated(Zone) & is_aggregated(Region),
    !is_aggregated(State)
  ) |> 
  model(
    STL(Nights)
  ) |> 
  components() |> 
  pivot_longer(
    cols = c(Nights, trend, season_year, remainder, season_adjust),
    names_to = "component",
    values_to = "value"
  ) |>
  filter(
    component == c("Nights", "trend"),
    year(Month) >= 2000
  ) |>
  as_tibble() |>
  mutate(component = recode(component,                     # give them nicer names
                            Nights = "Observed Nights",
                            trend  = "STL Decomposed Trend")) |> 
  ggplot(aes(Month, value, color = as.character(State))) +
    geom_line() +
    facet_wrap(~ component, ncol=1, scales = "free_y") +
    labs(
      x = "Month",
      y = "Visitor Nights (in thousands)",
      color = "State"
    ) +
    theme_bw()
```

Surprisingly, however, MinT underperforms OLS at higher aggregation level and even the base forecasts at country-wise. This contradicts with the canonical results from @Wickramasuriya2019-xq, prompting further investigation. Applying STL decomposition by @Bandara2022-og to state-aggregated observations reveals a structural change around 2016: four large states (New South Wales, Victoria, Queensland, Western Australia) develop a modest upward trend in visitor nights, as shown in @fig-STL-tourism. The trend is likely due to the increasing popularity of tourism in these places, as well as the impact of various factors such as economic growth, population growth, and changes in consumer preferences.

```{r}
#| label: fig-tourism-results-ex2016
#| fig-cap: "Relative improvement of the MSE of reconciled forecasts over the base forecasts for the Australian domestic tourism flows, excluding the period after 2016"
#| fig.width: 8
#| fig.height: 10
#| out.width: "100%"

tourism_MSE_ex2016 <- readRDS("literature/lib/tourism_MSE_ex2016.rds")

# % improv plot by level
tourism_MSE_ex2016 |>
  group_by(series, h) |>
  mutate(
    base_MSE = MSE[.model == "base"]
  ) |>
  ungroup() |>
  group_by(.model, level, h) |>
  summarise(
    MSE = mean(MSE),
    base_MSE = mean(base_MSE),
    pct_change = (MSE - base_MSE) / base_MSE * 100
  ) |>
  filter(.model != "mint_sample") |>
  # Reorder the levels according to the desired order
  mutate(
    level = factor(
      level,
      levels = desired_order,
      labels = str_replace_all(desired_order, "_", " ")  # replace "_" with " "
    )
  ) |>
  ggplot(aes(x = h, y = pct_change, color = .model)) +
    geom_line() +
    labs(x = "Horizon", y = "% improvements") +
    facet_wrap(~ level, ncol=2, scales = "free_y") +
    theme_bw() +
    scale_x_continuous(limits = c(1, 12), expand = c(0, 0)) +
    theme(legend.position = "bottom", legend.title = element_blank())
```

Neither ARIMA nor reconciliation strategies capture this shift, degrading accuracy at the most aggregated level. From the results in @fig-tourism-results-ex2016, excluding the post-2016 period (i.e., restricting the final training window to end on December 2015) restores reconciliation’s performance: both MinT methods now clearly outperform OLS and the base forecasts at all levels. Additionally, we notice that mint_n increasingly outperforms mint_shr at higher levels (Australia, State, and Australia by Purposes levels). This is an advantage of NOVELIST estimator. At higher levels most series are combinations of many bottom cells, their cross-series base forecast error correlations are strong and genuinely useful. NOVELIST keeps these strong links at those higher levels, while its thresholding feature minimises the effects of weak, noisy ones.

The empirical analysis on real, hierarchical Australian domestic tourism shreds lights on the strengths of the NOVELIST estimator--advantages that were not fully apparent in previous simulation studies. Going forward, we plan to embed this structure into our data-generating processes and to pinpoint contexts where NOVELIST’s thresholding yields gains beyond those of Shrinkage. Moreover, our findings underscore the need to accommodate structural changes--the post-2016 Australian domestic tourism demand shifts we encountered.

<!-- perhaps via robust trend adjustment or state-space extensions that jointly model coherence and temporal dynamics. -->

@@

@@

    "Final writing, presenation, submission"
  ),
  Deliverable = c(
    "Initial literature review + paper draft",
    "ReconCov package: Main codes for estimators + simulation framework + cross-validation",
    "NOVELIST initial simulation results + research proposal + presentation",
    "More simulation results on NOVELIST",
    "Empirical results on forecasting Tourism data",
    "New simulation design, POET & PC-adjusted NOVELIST simulation results",
    "Other estimators simulation results and full-fledged covariance estimators assessment",
    "Presentation and thesis manuscript"
  )
) |>
  kable(
    format = "latex",
    booktabs = TRUE,
    longtable = TRUE,
    caption = "Project Timeline and Key Milestones",
    escape = TRUE
  ) |>
  column_spec(1, width = "1.9cm") |>
  column_spec(2, width = "8cm") |>
  column_spec(3, width = "4.8cm") |>
  kable_styling(latex_options = "hold_position", font_size = 10)
```

>>>>

---
title: "Enhancing Forecast Reconciliation: A Study of Alternative Covariance Estimators"
author: 
  - Vincent Su
  - Shanika Wickramasuriya (supv.)
  - George Athanasopoulos (supv.)
format: 
  pdf:
    include-in-header:
      text: |
        <!-- \usepackage{amsmath} -->
        <!-- \usepackage{mathspec} -->
        \usepackage{algorithm}
        \usepackage{algpseudocode}
        \usepackage[colorlinks=true, linkcolor=blue, citecolor=blue, urlcolor=blue]{hyperref}
        \usepackage{array}
    mathspec: true
    number-sections: true
# crossref:
#   custom:
#     - kind: float
#       key: alg
#       reference-prefix: Algorithm
#       caption-prefix: Algorithm
#       # for PDF output only, tell LaTeX to use the `algorithm` env:
#       latex-env: algorithm
link-citations: true
execute:
  cache: true
  echo: false
  warning: false
bibliography: D:/Github/Recon_Honours_Thesis/references.bib
# bibliographystyle: apa
csl: D:/Github/Recon_Honours_Thesis/apa.csl
editor_options: 
  chunk_output_type: console
---

@@

```{r library}
library(ggplot2)
library(patchwork)
library(dplyr)
library(tidyr)
library(fable)
library(fabletools)
library(feasts)
library(tsibble)
library(lubridate)
library(knitr)
library(kableExtra)
library(stringr)
library(forcats)

@@

# Abstract {.unnumbered}

@@

<!-- -   Talk about description of the hierarchical time series data set (e.g., economic, financial, or synthetic data). -->

<!-- -   Characteristics such as dimensionality, frequency, and hierarchical structure. -->

@@

<!-- -   Design different cases of simulation studies and real-data experiments. -->

<!-- -   Metrics: Forecast accuracy (e.g., RMSE, MAE), reconciliation error reduction, and matrix stability... -->

<!-- -   Visualisation... -->

@@

<!-- Rather than collapsing each group into a single aggregate as in previous practice, we impose a more intricate, multi‐level aggregation path to stress‐test reconciliation methods. We first form ten intermediate series by summing bottom‐level series in consecutive blocks of ten. These ten series are then assigned to three second‐level aggregates--groups of four, three, and four--and finally summed to produce a single top‐level series. This asymmetric hierarchy creates overlapping correlation patterns: some level‐2 series share bottom‐level groups, while others draw from both. -->

<!-- emulating practical scenarios such as regional sales aggregations that span multiple product categories or overlapping territories. -->

@@

```{r}
tourism_plots <- readRDS("literature/tourism_plots.rds")
```

@@

The data are monthly time series spanning from January 1998 to December 2016, resulting in 228 observations per series, producing a canonical "$n \ll p$" setting which is ideal for evaluating reconciliation approaches that rely on high-dimensional covariance estimation. The extreme dimensionality over sample size mirrors many contemporary business problems, for instance, Starbucks drink sales. Tourism demand is also economically vital yet highly volatile, with geographical and purpose‑specific patterns create a realistic stress‑test for reconciliation algorithms.

<!-- Consequently, this panel offers both a rich policy case study and a stringent statistical laboratory for comparing reconciliation strategies that exploit cross-series information to improve forecasts when historical data are scarce. -->

@@

@Wickramasuriya2019-xq also argued that modelling spatial autocorrelations directly from the start would be challenging as in this case of a large collection of time series. Post-processing reconciliation approaches have the advantage to implicitly model this spatial autocorrelation structure, especially true for MinT.

To assess forecasting performance, we adopt a rolling-window cross-validation scheme. Beginning with the first 120 monthly observations (January 1998-December 2005) as the initial training set, we obtain the best-fitted ARIMA model for each of the 560 series via the automatic algorithm by minimising AICc from @Hyndman2008-rf. The 1- to 12-step-ahead base forecasts are then generated by these ARIMA models, and then reconciled using multiple approaches. We then roll the training window forward by one month and refit all models, rebuild reconciliations, and produce another batch of 1- to 12-step-ahead forecasts, repeating until the training set reaches December 2015. In total, this results in 133 out-of-sample windows.

<!-- 
Should MinT-sample be included in the paper?

Note that the number of series is larger than the number of observations (560 compared to 96), hence the sample covariance matrix is not positive definite and will not be considered.
-->

```{r}
#| label: fig-tourism-MSE-line
#| fig.height: 6
#| fig.width: 8
#| fig-cap: "Percentage relative improvement in the mean squared error (MSE) of different reconciled forecasts over the base forecasts for the Australian domestic tourism data, for 1 to 12 steps ahead forecasts. The positive entries indicate an decrease in MSE."
tourism_plots$MSE_line + 
  labs(title = NULL, subtitle = NULL) + 
  theme(
    legend.position = "bottom", 
    legend.direction = "vertical", 
    legend.title.position = "left"
  ) + 
  scale_x_continuous(breaks = c(1:12), minor_breaks = NULL)
```

```{r}
#| label: fig-tourism-MSE-heat
#| fig.height: 5
#| fig.width: 7
#| out.width: "90%"
#| fig-cap: "Heatmap of relative improvement in the mean squared error (MSE) of different reconciled forecasts over the base forecasts for the Australian domestic tourism data, for 1 to 12 steps ahead forecasts. The values are scaled to the range of 0 to 100 for better visualisation, with darker colors indicating greater improvement and best performance is noted by a star."

tourism_plots$MSE_heat + 
  labs(title = NULL, subtitle = NULL) + 
  theme(legend.title = element_blank(), legend.position = "bottom") + 
  scale_x_continuous(breaks = c(1:12), minor_breaks = NULL)
```

@fig-tourism-MSE-line and @fig-tourism-MSE-heat show that reconciliation is beneficial for point forecasts: across horizons $h=1,2,\dots,12$, all reconciled methods improve MSE over incoherent base ARIMA on average. Among vanilla MinT variants, the shrinkage estimator (*MinT-S*) marginally outperforms NOVELIST (*MinT-N*) at most horizons, though the gap is small. Adjusting for a single dominant factor via PC decomposition tightens performance further: both *MinT-S(PC1)* and *MinT-N(PC1)* beat their unadjusted counterparts, with *MinT-N(PC1)* narrowly ahead. Adding more than one PC brings no additional benefit, likely due to injecting estimation noise from weaker components. Variants that modify the multi-step covariance, either via scaled-variance or direct h-step residual covariances, underperform standard MinT, suggesting that extra estimation at horizon $h>1$ is not rewarded in this setting.

```{r}
#| label: fig-tourism-probscores
#| fig.height: 12
#| fig.width: 12
#| fig-cap: "Percentage relative improvement in the Winkler score at 80% and 95% nominal coverage, CRPS, and Energy score of multiple reconciled forecasts over the base forecasts for the Australian domestic tourism data, for 1-step-ahead forecasts. The positive entries indicate an decrease in the probabilistic scores."

# a 2x2 showing 4 plots showing the 1-step-ahead % improv in Winkler score 80, 95, CRPS, and Energy

tourism_plots$W80 + labs(title = "Winkler score 80%", subtitle = NULL) + 
  tourism_plots$W95 + labs(title = "Winkler score 95%", subtitle = NULL) +
  tourism_plots$CRPS + labs(title = "CRPS", subtitle = NULL) +
  tourism_plots$Energy + labs(title = "Energy score", subtitle = NULL) +
  plot_layout(ncol = 2, guides = "collect", axis_titles = "collect") & 
  theme(legend.position = "bottom", legend.title = element_blank())

```

Turning to probabilistic forecasts (1-step-ahead forecasts), @fig-tourism-probscores shows that *MinT-N* consistently outperforms *MinT-S* across CRPS and both Winkler scores, and PC1-adjustment again yields the best improvements. Results are aligned across probabilistic scoring rules, but the 95% Winkler reveals a notable pattern, as all MinT variants lose to the base while *OLS* performs best.

\color{red}
This points to tail-calibration sensitivity at high nominal coverage. The reconciliation that helps central accuracy can degrade extreme-tail interval performance when uncertainty is misallocated across the hierarchy.
\color{black}

In the multivariate evaluation, the Energy score places *OLS* close to the PC1-adjusted MinT methods, indicating that simple coherent projection might be able to capture a substantial share of cross-series dependence benefits even without sophisticated covariance regularisation.

The summary radar graph in @fig-tourism-radar consolidates these findings: among the selected MinT models by probabilistic criteria, *MinT-N(PC1)* clearly leads across CRPS, W80, W95, and Energy, extending the improvements seen in the single-metric panels.

Taken together, the evidence supports the use of reconciliation for both point and probabilistic forecasts in this high-dimensional setting. For point forecasts, MinT with shrinkage is a solid default; for probabilistic forecasts, NOVELIST performs more reliably. PC adjustment with a single dominant factor consistently enhances performance and should be considered when a dominant latent factor is evident. More complex adjustments, such as multiple PCs or horizon-specific covariances, add estimation noise without clear benefits and can be omitted for parsimony.

```{r}
#| label: fig-tourism-radar
#| fig.height: 8
#| fig.width: 8
#| out.width: "90%"
#| fig-cap: "Radar plot of relative improvements in probabilistic scores (Winkler 80, Winkler 95, CRPS, Energy) over the base forecasts. The scores are scaled to a range of 0 to 100, with larger values indicating better performance. Only the top 5 models are shown."

# a radar plot of 5 top models based on probabilistic scores (Winkler 80, 95, CRPS, Energy) scaled to 0-100 
par(mfrow=c(1,1), mar=c(0,0,0,0))
tourism_plots$radar
```

@@

<!-- 
\pagebreak

@@

    "Final writing, presenation, submission"
  ),
  Deliverable = c(
    "Initial literature review + paper draft",
    "ReconCov package: Main codes for estimators + simulation framework + cross-validation",
    "NOVELIST initial simulation results + research proposal + presentation",
    "More simulation results on NOVELIST",
    "Empirical results on forecasting Tourism data",
    "New simulation design, POET & PC-adjusted NOVELIST simulation results",
    "Other estimators simulation results and full-fledged covariance estimators assessment",
    "Presentation and thesis manuscript"
  )
) |>
  kable(
    format = "latex",
    booktabs = TRUE,
    longtable = TRUE,
    caption = "Project Timeline and Key Milestones",
    escape = TRUE
  ) |>
  column_spec(1, width = "1.9cm") |>
  column_spec(2, width = "8cm") |>
  column_spec(3, width = "4.8cm") |>
  kable_styling(latex_options = "hold_position", font_size = 10)
```
-->
